---
title: "Actinobacteria: an overlooked source of soil aggregate stability?"
author: "Reed Close, Kinsey"
date: "2025-11-19"
output: html_document
---

Analysis and supplementary materials for the manuscript "Bacteria: an overlooked source of soil aggregate stability?":

Reed Close K, LeMaster D, Schartiger R, Guthrie K, Kane J, Kotcon J, Morrissey EM. 2025. Bacteria: an overlooked source of soil aggregate stability?"

Corresponding Author: Kinsey Close Reed
School of Natural Resources and the Environment, Davis College of Agriculture and Natural Resources, West Virginia University, Morgantown, WV, USA
kyreed@mail.wvu.edu
ORCID ID: 0000-0001-5517-8664


Install and load required packages
```{r env set up, include=FALSE}

#Clear Environment
rm(list=ls()) 

# ——— Install required packages if not already installed ———
required_packages <- c(
  "ARTool", "car", "caret", "compositions", "DHARMa", "EnvStats",
  "emmeans", "esquisse", "elevatr", "ggridges", "ggdist", "ggmap",
  "ggpubr", "ggrepel", "ggspatial", "glmnet", "janitor", "performance",
  "patchwork", "raster", "rstatix", "sf", "svglite", "tigris",
  "tidyverse", "viridis"
)
new_packages <- setdiff(required_packages, rownames(installed.packages()))
if (length(new_packages) > 0) {
  install.packages(new_packages)
}



# ——— Load required packages ———
library(ARTool)      # ANOVA-type models
library(car)         # Companion to Applied Regression
library(caret)       # Repeated cross-validation, model training
library(compositions) # Compositional data analysis
library(DHARMa)      # Residual diagnostics for mixed models
library(EnvStats)    # Environmental & ecological statistics
library(emmeans)     # Estimated marginal means (post-hoc)
library(esquisse)    # ggplot2 GUI
library(elevatr)     # Elevation data via web APIs
library(ggridges)    # Ridgeline plots
library(ggdist)      # Visualizations of distributions
library(ggmap)       # Maps in ggplot2
library(ggpubr)      # ‘ggplot2’ based publication ready plots
library(ggrepel)     # Smart text repelling for ggplot2
library(ggspatial)   # Spatial layers in ggplot2
library(glmnet)      # Lasso & ridge regression
library(janitor)     # Data cleaning helpers
library(performance) # Model performance checks
library(patchwork)   # Compose ggplots
library(raster)      # Raster data handling
library(rstatix)     # Pipe-friendly statistical tests
library(sf)          # Simple features for spatial data
library(svglite)     # SVG graphics device
library(tigris)      # Census TIGER/Line shapefiles
library(tidyverse)   # Metapackage: dplyr, tidyr, ggplot2, forcats, etc.
library(viridis)     # Color scales


#Set Dplyr's Select Function as Default
select<-dplyr::select
```


Data preparation
```{r dataprep}
data <-read_csv("data.csv", col_names=TRUE, trim_ws=TRUE)
site_info <- read_csv("site_information.csv", col_names=TRUE, trim_ws=TRUE)
#Q in 2022 only
#make integers numeric 
data <- data %>% mutate_if(is.integer, as.numeric)
data$year <- as.factor(data$year)



#Create Slaking Index called "stability" so get inverse (for reader clarity) and workable distribution

data$stability <-log(data$slaking_index)*-1
hist(data$stability)
densityplot(data$stability)

#will not remove remove Q from 2022 because looking at each year independently 

#Euk and fungi log tranform
densityplot((data$eukaryote))
data$log_eukaryote <- log(data$eukaryote)
densityplot((data$log_eukaryote))

densityplot((data$fungi))
data$log_fungi <- log(data$fungi)
densityplot((data$log_fungi))


#Subset by year

data_2022 <- subset(data, year == 2022)
data_2023 <- subset(data, year == 2023)

```

Supplemental Figure 1 - Site Location Map
```{r SF1_map}

# Retrieve US states (cartographic boundary version)
wv <- states(cb = TRUE) %>%
  filter(NAME == "West Virginia") %>%
  st_transform(crs = 4326)  # WGS84 lat/lon

# Data Frame with sample site coords
#not provided to public for farmer privacy. Please reach out to the author if coordinates are requested.
fig1 <- site_info
samplesites <- fig1
#note RF has 3 sites
#SL not included. But is filtered out of dataset anyway
#HH included when had been removed from dataset in some instances? 

samplesites
# Bounding box for WV
wv_bbox <- st_bbox(wv)

# Download elevation data (~90m resolution SRTM)
elev <- get_elev_raster(locations = wv, z = 8, clip = "locations")

# Convert to dataframe for ggplot
elev_df <- as.data.frame(elev, xy = TRUE)
names(elev_df)[3] <- "elevation"


# Convert to sf object
sites_sf <- st_as_sf(samplesites, coords = c("longitude", "latitude"), crs = 4326)
#crs = 4326 sets the coordinate reference system to WGS84 lat/lon (EPSG:4326).
#coordinates not provided in this dataset, if needed please reach out to corresponding author

# Create a new column 'site_color' that is "red" for A, B, D and "black" otherwise.
sites_sf <- sites_sf %>%
  mutate(site_color = if_else(FarmID %in% c("A", "B", "D"), "red", "black"))
  
sites_sf$site_color[sites_sf$FarmID == "Q"] <- "blue" #only in 2022 data set


ggplot() +
  # WV boundary polygon
  geom_sf(data = wv, fill = "white", color = "black") +
  
  # Sample sites
  geom_sf(data = sites_sf, aes(color = site_color), size = 3) + # aes(color = FarmID)
  # Use the identity scale so the color values are taken lit
   scale_color_identity() +
  # Optional: scale bar & north arrow
  annotation_scale(location = "bl", width_hint = 0.4) +  # scale bar in bottom-left
  annotation_north_arrow(location = "bl", which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  
  # Labels & theme
  labs(title = "Sample Site Locations in West Virginia, USA",
       color = "Site ID") +
  theme_minimal(base_size = 14, base_family = "Helvetica") +
  theme(
    panel.grid.major = element_blank(),
    legend.position = "right",
    plot.title = element_text(face = "bold", hjust = 0.5)
  )


# Use WGS84 projection
wv_wgs84 <- st_transform(wv, crs = 4326)

# Download elevation (z = 8 is a good resolution for WV)
elev <- get_elev_raster(locations = wv_wgs84, z = 8, clip = "locations")

# Generate elevation contour lines every 200 meters (adjust as needed)
contours <- rasterToContour(elev, levels = seq(0, 1600, by = 100))  # in meters
contours_sf <- st_as_sf(contours)

ggplot() +
  # Elevation contours
  geom_sf(data = contours_sf, aes(), color = "grey50", size = 0.3) +
  
  # WV boundary
  geom_sf(data = wv_wgs84, fill = NA, color = "black", size = 0.8) +
  
  # Sample sites
  geom_sf(data = sites_sf, aes(color = site_color), size = 3) +
  scale_color_identity() +
  
  # North arrow and scale
  annotation_scale(location = "bl", width_hint = 0.4) +
  annotation_north_arrow(location = "bl", which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  
  # Labels & theme
  labs(title = "Sample Site Locations",
      # subtitle = "Contours derived from SRTM elevation (every 100 m)",
       color = "Site ID") +
  theme_minimal(base_size = 14, base_family = "Helvetica") +
  theme(
    panel.grid.major = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )

ggsave("sitemap_elevation_contours.png")


# Make sure sites_sf and elev have matching CRS
sites_sf_proj <- st_transform(sites_sf, crs = crs(elev))

# Extract elevation values for each point
elev_values <- raster::extract(elev, sites_sf_proj)

# Add elevation as a column to your sites data
sites_sf$Elevation_m <- elev_values
#write.csv(sites_sf, "sites_sf_elev.csv")


ggsave("wv_sample_sites.png", width = 8, height = 6, dpi = 600)



```


Figure 2 - PLFA Biomass x soil stability scatterplots
```{r F2}

#removing outliers
#tested for outliers further below by using simulateResiduals(fittedModel = m_simp22) and check_model(m_simp22)
#came back up here to remove from whole dataset where relevant

#L 2022 one replicate is an outlier for Act, GP, AMF, fungi (log)
#I'd say safe to remove from 2022 dataset altogether
#Others-
#2022: D one replicate (gram_negative), A one replicate (Euk, super high but within trend)
#2023: F one replicate (log_Fun), eukaryote has a few high values, but seems consistent and stays within trend? 

#For mult. regression, remove for each where relevant and use drop.na()
#For PCA, only needed to remove outlier for AMF, Act (only DS-L 22)
#For Boxplot, make individual obs NA? 
#For Ridge, make individual obs NA and drop.na()

data_filt <- data %>%
  filter(index != 6)  #removes one L rep 22 and 23

#outliers in log_Euk 2023 

densityplot(data_2023$log_eukaryote)
#Clearly identifies whether the 3 largest points significantly deviate from the distribution.
rosnerTest(data_2023$log_eukaryote, k = 3) #two of them 
#i   Mean.i      SD.i     Value Obs.Num    R.i+1 lambda.i+1 Outlier
#1 0 1.238579 0.4964266 2.8732256       9 3.292826   3.179900    TRUE
#2 1 1.209389 0.4488535 2.6847133      35 3.286872   3.173022    TRUE
#3 2 1.182565 0.4051672 0.1939207      13 2.440091   3.165989   FALSE


# Replace a single outlier values with NA
data_filt$gram_negative[data_filt$index == "28"] <- NA #D 22
data_filt$log_eukaryote[data_filt$index == "33"] <- NA #A 22
data_filt$eukaryote[data_filt$index == "33"] <- NA #A 22
data_filt$log_eukaryote[data_filt$index == "69"] <- NA #I 23
data_filt$log_eukaryote[data_filt$index == "95"] <- NA #B 23


data_filt$log_fungi[data_filt$index == "86"] <- NA #QF-H 23
data_filt$fungi[data_filt$index == "86"] <- NA #QF-H 23

#all others seem good after removing above
densityplot(data_2023$actinomycetes, na.rm = TRUE)

densityplot(data_2023$log_fungi, na.rm = TRUE)
rosnerTest(data_2023$actinomycetes, k = 1)

densityplot(data_2023$actinomycetes, na.rm = TRUE)
densityplot(data_2023$am_fungi, na.rm = TRUE)
densityplot(data_2023$gram_positive, na.rm = TRUE)
densityplot(data_2023$gram_negative, na.rm = TRUE)

#no outliers in any
densityplot(data_2022$log_eukaryote, na.rm = TRUE)
rosnerTest(data_2022$log_eukaryote, k = 3)
densityplot(data_2022$actinomycetes, na.rm = TRUE)
rosnerTest(data_2022$actinomycetes, k = 1)
densityplot(data_2022$am_fungi, na.rm = TRUE)
rosnerTest(data_2022$am_fungi, k = 1)
densityplot(data_2022$gram_positive, na.rm = TRUE)
rosnerTest(data_2022$gram_positive, k = 1)

densityplot(data_2022$gram_negative, na.rm = TRUE)
densityplot(data_2022$log_fungi, na.rm = TRUE)


#redo by year

data_2022 <- subset(data_filt, year == 2022)
data_2023 <- subset(data_filt, year == 2023)

#Scatterplots are colored by year, and show sep Rsq for each year.
#reg lm with only the one group

#Actinobacteria 
densityplot(data_2022$actinomycetes)
m_simp22 <- lm(stability ~ actinomycetes, data = data_2022)
summary(m_simp22) #p = ****
 r2_22 <- summary(m_simp22)$adj.r.squared 
 r2_22

 simulationOutput22<- simulateResiduals(fittedModel = m_simp22)
plot(simulationOutput22) 
check_model(m_simp22) #DS-L was outlier for 2022 Actino

m_simp23 <- lm(stability ~ actinomycetes, data = data_2023)
summary(m_simp23) #p = ****
r2_23 <- summary(m_simp23)$adj.r.squared
r2_23
simulationOutput23<- simulateResiduals(fittedModel = m_simp23)
plot(simulationOutput23)  
check_model(m_simp23)


# 2. Create labels for each R²
label_spring <- paste0("2022 Adj.R² = ", round(r2_22, 2))
label_fall   <- paste0("2023 Adj.R² = ", round(r2_23, 2))


plot_act <- ggplot(data = data_filt, aes(x = actinomycetes, y = stability, color = as.factor(year), size = om_perc)) +
  geom_point(alpha = 0.8) +
  geom_smooth(data = data_filt,
              aes(x = actinomycetes, y = stability, color = as.factor(year)),
              method = "lm",
              se = FALSE,
              inherit.aes = FALSE) +
   
  scale_size_continuous(range = c(1.5, 4.5)) +
  scale_color_manual(values = c("2022" = "#1f77b4", "2023" = "darkorange")) +
  labs(x = "Actinobacteria (nmol per g soil)", y = "Macroaggregate Stability Index",color = element_blank(), size = "% Org.Matter") +
  
  # 4. Annotate R² values
  annotate("text", x = 13, y = 2.3, label = label_spring, color = "#1f77b4", size = 3.5) +
  annotate("text", x = 13, y = 2.1, label = label_fall,   color = "darkorange", size = 3.5) +
  theme_classic(base_size = 10, base_family = "Helvetica") +
  theme(
    # Remove minor grid lines
    panel.grid.minor = element_blank(),
    # Black axis lines
    axis.line = element_line(color = "black", size = 0.5),
    # Ticks inside, slightly smaller
    axis.ticks = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(0.15, "cm"),
    # Control margin around the plot
    plot.margin = margin(5, 5, 5, 5),
    # Legend
    legend.position = "right",
    legend.background = element_blank()
  )
plot_act


#AMF
m_amf_simp22 <- lm(stability ~ am_fungi, data = data_2022)
summary(m_amf_simp22) #p = ****
amf_r2_22 <- summary(m_amf_simp22)$adj.r.squared
amf_r2_22
simulationOutput22<- simulateResiduals(fittedModel = m_amf_simp22)
plot(simulationOutput22) 
check_model(m_amf_simp22) #DS-L outlier

m_amf_simp23 <- lm(stability ~ am_fungi, data = data_2023)
summary(m_amf_simp23) #p = ****
amf_r2_23 <- summary(m_amf_simp23)$adj.r.squared 
amf_r2_23
simulationOutput<- simulateResiduals(fittedModel = m_amf_simp23)
plot(simulationOutput) 
check_model(m_amf_simp23) 


# 2. Create labels for each R²
label_spring_amf <- paste0("2022 Adj.R² = ", round(amf_r2_22, 2))
label_fall_amf   <- paste0("2023 Adj.R² = ", round(amf_r2_23, 2))


plot_amf <- ggplot(data = data_filt, aes(x = am_fungi, y = stability, color = as.factor(year), size = om_perc)) +
  geom_point(alpha = 0.8) +
  geom_smooth(data = data_filt,
              aes(x = am_fungi, y = stability, color = as.factor(year)),
              method = "lm",
              se = FALSE,
              inherit.aes = FALSE) +
  scale_size_continuous(range = c(1.5, 4.5)) +
  scale_color_manual(values = c("2022" = "#1f77b4", "2023" = "darkorange")) +
  labs(x = "AM Fungi (nmol per g soil)", y = "Macroaggregate Stability Index",color = element_blank(), size = "% Org.Matter") +
  
  # 4. Annotate R² values
  annotate("text", x = 4, y = 2.3, label = label_spring_amf, color = "#1f77b4", size = 3.5) +
  annotate("text", x = 4, y = 2.1, label = label_fall_amf,   color = "darkorange", size = 3.5) +

  theme_classic(base_size = 10, base_family = "Helvetica") +
  theme(
    # Remove minor grid lines
    panel.grid.minor = element_blank(),
    # Black axis lines
    axis.line = element_line(color = "black", size = 0.5),
    # Ticks inside, slightly smaller
    axis.ticks = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(0.15, "cm"),
    # Control margin around the plot
    plot.margin = margin(5, 5, 5, 5),
    # Legend
    legend.position = "right",
    legend.background = element_blank()
  )
plot_amf


#Gram Negative

m_gn_simp22 <- lm(stability ~ gram_negative, data = data_2022) #same with or without na.action=na.exclude
summary(m_gn_simp22) #p = ****
gn_r2_22 <- summary(m_gn_simp22)$adj.r.squared 
gn_r2_22
simulationOutput22<- simulateResiduals(fittedModel = m_gn_simp22)
plot(simulationOutput22) 
check_model(m_gn_simp22) #RF1-C outlier


m_gn_simp23 <- lm(stability ~ gram_negative, data = data_2023)
summary(m_gn_simp23) #p = ****
gn_r2_23 <- summary(m_gn_simp23)$adj.r.squared 
gn_r2_23
simulationOutput <- simulateResiduals(fittedModel = m_gn_simp23)
plot(simulationOutput) 
check_model(m_gn_simp23) 

# 2. Create labels for each R²
gn_label_spring <- paste0("2022 Adj.R² = ", round(gn_r2_22, 2))
gn_label_fall   <- paste0("2023 Adj.R² = ", round(gn_r2_23, 2))


plot_gn <- ggplot(data = data_filt, aes(x = gram_negative, y = stability, color = as.factor(year), size = om_perc)) +
  geom_point(alpha = 0.8) +
  geom_smooth(data = data_filt,
              aes(x = gram_negative, y = stability, color = as.factor(year)),
              method = "lm",
              se = FALSE,
              inherit.aes = FALSE) +
  scale_size_continuous(range = c(1.5, 4.5)) +
  scale_color_manual(values = c("2022" = "#1f77b4", "2023" = "darkorange")) +
  labs(x = "Gram-Negative Bacteria (nmol per g soil)", y = "Macroaggregate Stability Index",color = element_blank(), size = "% Org.Matter") +
  
  # 4. Annotate R² values
  annotate("text", x = 35, y = 2.3, label = gn_label_spring, color = "#1f77b4", size = 3.5) +
  annotate("text", x = 35, y = 2.1, label = gn_label_fall,   color = "darkorange", size = 3.5) +
  
  theme_classic(base_size = 10, base_family = "Helvetica") +
  theme(
    # Remove minor grid lines
    panel.grid.minor = element_blank(),
    # Black axis lines
    axis.line = element_line(color = "black", size = 0.5),
    # Ticks inside, slightly smaller
    axis.ticks = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(0.15, "cm"),
    # Control margin around the plot
    plot.margin = margin(5, 5, 5, 5),
    # Legend
    legend.position = "right",
    legend.background = element_blank()
  )
plot_gn

#Gram Positive

m_gp_simp22 <- lm(stability ~ gram_positive, data = data_2022)
summary(m_gp_simp22) #p = ****
gp_r2_22 <- summary(m_gp_simp22)$adj.r.squared 
gp_r2_22 
simulationOutput22<- simulateResiduals(fittedModel = m_gp_simp22)
plot(simulationOutput22) 
check_model(m_gp_simp22) #DS-L outlier


m_gp_simp23 <- lm(stability ~ gram_positive, data = data_2023)
summary(m_gp_simp23) #p = ****
gp_r2_23 <- summary(m_gp_simp23)$adj.r.squared 
gp_r2_23
simulationOutput<- simulateResiduals(fittedModel = m_gp_simp23)
plot(simulationOutput) 
check_model(m_gp_simp23) 


# 2. Create labels for each R²
gp_label_spring <- paste0("2022 Adj.R² = ", round(gp_r2_22, 2))
gp_label_fall   <- paste0("2023 Adj.R² = ", round(gp_r2_23, 2))

plot_gp <- ggplot(data = data_filt, aes(x = gram_positive, y = stability, color = as.factor(year), size = om_perc)) +
  geom_point(alpha = 0.8) +
  geom_smooth(data = data_filt,
              aes(x = gram_positive, y = stability, color = as.factor(year)),
              method = "lm",
              se = FALSE,
              inherit.aes = FALSE) +
  scale_size_continuous(range = c(1.5, 4.5)) +
  scale_color_manual(values = c("2022" = "#1f77b4", "2023" = "darkorange")) +
  labs(x = "Other Gram-Positive Bacteria (nmol per g soil)", y = "Macroaggregate Stability Index",color = element_blank(), size = "% Org.Matter") +
  
  # 4. Annotate R² values
  annotate("text", x = 22, y = 2.3, label = gp_label_spring, color = "#1f77b4", size = 3.5) +
  annotate("text", x = 22, y = 2.1, label = gp_label_fall,   color = "darkorange", size = 3.5) +
  theme_classic(base_size = 10, base_family = "Helvetica") +
  theme(
    # Remove minor grid lines
    panel.grid.minor = element_blank(),
    # Black axis lines
    axis.line = element_line(color = "black", size = 0.5),
    # Ticks inside, slightly smaller
    axis.ticks = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(0.15, "cm"),
    # Control margin around the plot
    plot.margin = margin(5, 5, 5, 5),
    # Legend
    legend.position = "right",
    legend.background = element_blank()
  )
plot_gp


#eukaryotes

m_euk_simp22 <- lm(stability ~ log_eukaryote, data = data_2022)
summary(m_euk_simp22) #p = **
euk_r2_22 <- summary(m_euk_simp22)$adj.r.squared
euk_r2_22
simulationOutput22<- simulateResiduals(fittedModel = m_euk_simp22)
plot(simulationOutput22) 
check_model(m_euk_simp22) #RF2-L super high 


m_euk_simp23 <- lm(stability ~ log_eukaryote, data = data_2023) #data_2023_euk
summary(m_euk_simp23) #p = **
euk_r2_23 <- summary(m_euk_simp23)$adj.r.squared 
euk_r2_23
simulationOutput<- simulateResiduals(fittedModel = m_euk_simp23)
plot(simulationOutput) #good
check_model(m_euk_simp23) 

# 2. Create labels for each R²
euk_label_spring <- paste0("2022 Adj.R² = ", round(euk_r2_22, 2))
euk_label_fall   <- paste0("2023 Adj.R² = ", round(euk_r2_23, 2))

plot_euk <- ggplot(data = data_filt, aes(x = log_eukaryote, y = stability, color = as.factor(year), size = om_perc)) +
  geom_point(alpha = 0.8) +
  geom_smooth(data = data_filt,
              aes(x = log_eukaryote, y = stability, color = as.factor(year)),
              method = "lm",
              se = FALSE,
              inherit.aes = FALSE) +
  scale_size_continuous(range = c(1.5, 4.5)) +
  scale_color_manual(values = c("2022" = "#1f77b4", "2023" = "darkorange")) +
  labs(x = "Other Eukaryotes log(nmol per g soil)", y = "Macroaggregate Stability Index",color = element_blank(), size = "% Org.Matter") +
  
  # 4. Annotate R² values
  annotate("text", x = 0.1, y = 2.2, label = euk_label_spring, color = "#1f77b4", size = 3.5) +
  annotate("text", x = 0.1, y = 2.0, label = euk_label_fall,   color = "darkorange", size = 3.5) +
  theme_classic(base_size = 10, base_family = "Helvetica") +
  theme(
    # Remove minor grid lines
    panel.grid.minor = element_blank(),
    # Black axis lines
    axis.line = element_line(color = "black", size = 0.5),
    # Ticks inside, slightly smaller
    axis.ticks = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(0.15, "cm"),
    # Control margin around the plot
    plot.margin = margin(5, 5, 5, 5),
    # Legend
    legend.position = "right",
    legend.background = element_blank()
  )
plot_euk


#fungi


m_fun_simp22 <- lm(stability ~ log_fungi, data = data_2022) #_fun
summary(m_fun_simp22) #p = NS
fun_r2_22 <- summary(m_fun_simp22)$adj.r.squared 
fun_r2_22
simulationOutput22<- simulateResiduals(fittedModel = m_fun_simp22)
plot(simulationOutput22) 
check_model(m_fun_simp22) #DS-L outlier 


m_fun_simp23 <- lm(stability ~ log_fungi, data = data_2023) 
summary(m_fun_simp23) #p = ** 0.008
fun_r2_23 <- summary(m_fun_simp23)$adj.r.squared 
fun_r2_23
simulationOutput<- simulateResiduals(fittedModel = m_fun_simp23)
plot(simulationOutput) 
check_model(m_fun_simp23) #QF-H outlier

# 2. Create labels for each R²
fun_label_spring <- paste0("2022 Adj.R² = ", round(fun_r2_22, 2))
fun_label_fall   <- paste0("2023 Adj.R² = ", round(fun_r2_23, 2))


fun_annotation <- annotation_custom(grid::textGrob(label = "example watermark",
                                               x = unit(0.75, "npc"), y = unit(0.25, "npc"),
                                               gp = grid::gpar(cex = 2)))


plot_fun <- ggplot(data = data_filt, aes(x = log_fungi, y = stability, color = as.factor(year), size = om_perc)) +
  geom_point(alpha = 0.8) +
  geom_smooth(data = data_filt,
              aes(x = log_fungi, y = stability, color = as.factor(year)),
              method = "lm",
              se = FALSE,
              inherit.aes = FALSE) +
  scale_size_continuous(range = c(1.5, 4.5)) +
  scale_color_manual(values = c("2022" = "#1f77b4", "2023" = "darkorange")) +
  labs(x = "Other Fungi log(nmol per g soil)", y = "Macroaggregate Stability Index",color = element_blank(), size = "% Org.Matter") +
  
  # 4. Annotate R² values
  annotate("text", x = 0.25, y = 2.2, label = fun_label_spring, color = "#1f77b4", size = 3.5) +
  annotate("text", x = 0.25, y = 2.0, label = fun_label_fall,   color = "darkorange", size = 3.5) +
  theme_classic(base_size = 10, base_family = "Helvetica") +
  theme(
    # Remove minor grid lines
    panel.grid.minor = element_blank(),
    # Black axis lines
    axis.line = element_line(color = "black", size = 0.5),
    # Ticks inside, slightly smaller
    axis.ticks = element_line(color = "black", size = 0.5),
    axis.ticks.length = unit(0.15, "cm"),
    # Control margin around the plot
    plot.margin = margin(5, 5, 5, 5),
    # Legend
    legend.position = "right",
    legend.background = element_blank()
  )
plot_fun




#all together
ggarrange(plot_act,
          plot_amf,
          plot_gn,
          plot_gp,
          plot_fun,
          plot_euk,
          ncol = 2, nrow = 3,
          labels = c("A", "B", "C", "D", "E", "F"),
          common.legend = TRUE, legend = "right")


ggsave("Fig2.png")




```

Supplemental Figure 2 - Soil Chemical and Biological Properties
```{r SF2}

selectedprops <- data_filt %>% 
  dplyr::select(anon_id,index, stability,
                om_perc,
                actinomycetes, am_fungi, 
                gram_negative, gram_positive, eukaryote, fungi,
                ph, cec_meq_100g,
                p_lb_a, mg_lb_a, ca_lb_a, k_lb_a,
                zn_ppm, b_ppm, cu_ppm, mn_ppm, fe_ppm,
                so4_ppm,
                year
  )

selectedprops <- selectedprops %>%
  group_by(anon_id, year,index) %>%
  mutate(total_plfa = sum(actinomycetes, am_fungi, gram_negative, gram_positive, eukaryote, fungi,na.rm = TRUE))

#move to where you want it in the graph layout
selectedprops<- selectedprops %>% relocate(total_plfa, .after = om_perc)

selectedprops$year <- as.factor(selectedprops$year)
selectedprops$anon_id <- as.factor(selectedprops$anon_id)

#remove index
selectedprops <- subset(selectedprops, select = -index)
# 4. pivot to long form
soil_vars <- setdiff(names(selectedprops), c("anon_id","year"))

# 3. average the three replicates per farm × year
summ_props <- selectedprops %>%
  group_by(anon_id, year) %>%
  summarize(across(all_of(soil_vars), mean, na.rm=TRUE), .groups="drop")

# 4. compute unique farm order by overall mean OM

om_order <- summ_props %>%
  group_by(anon_id) %>%
  summarize(mean_om = mean(om_perc, na.rm=TRUE), .groups="drop") %>%
  arrange(desc(mean_om)) %>%
  pull(anon_id)

# 5. pivot & factor
long_props <- summ_props %>%
  # enforce farm ordering, then recode to letters
  mutate(
    id_ordered = factor(anon_id, levels = om_order),
    year    = factor(year,    levels = c("2022", "2023"))
  ) %>%
  pivot_longer(
    cols      = all_of(soil_vars),
    names_to  = "Variable",
    values_to = "Value"
  ) %>%
  # preserve the original soil‐var sequence
  mutate(Variable = factor(Variable, levels = soil_vars))

# 5. named labels for each facet
var_labels <- c(
  om_perc = "Organic matter (%)", ph       = "pH",
  stability = "Macroaggregate stability index",         	
  cec_meq_100g  = "CEC (mEq 100g)",
  p_h = "ph",
  p_lb_a   = "P (lb/Ac)",       mg_lb_a  = "Mg (lb/Ac)",
  ca_lb_a  = "Ca (lb/Ac)",       k_lb_a   = "K (lb/Ac)",
  zn_ppm   = "Zn (ppm)",          b_ppm    = "B (ppm)",
  cu_ppm   = "Cu (ppm)",          mn_ppm   = "Mn (ppm)",
  fe_ppm   = "Fe (ppm)",          so4_ppm  = "SO4 (ppm)",
  total_plfa_nmol_p_g_soil   = "Total PLFAs (nmol per g soil)",
  actinomycetes     = "Actinobacteria (nmol per g soil)",  am_fungi = "AM Fungi (nmol per g soil)",
  gram_positive      = "Gram-positive (nmol per g soil)", gram_negative   = "Gram-negative (nmol per g soil)",
  eukaryote    = "Eukaryotes (nmol per g soil)",    fungi  = "Fungi (nmol per g soil)"
)


#plot
ggplot(long_props, aes(x = anon_id, Value, color=year, group=year)) +
  geom_line(size=0.8) +
  geom_point(size=2) +
  facet_wrap(~ Variable, scales="free_y", ncol=3,
             labeller = labeller(Variable = var_labels)) +
  scale_color_manual(values = c("2022"="#1f77b4","2023"="darkorange")) +
  theme_bw(base_size=11,base_family = "Helvetica") +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    strip.text     = element_text(face="bold", size=9)
  ) +
  labs(
    x     = "Site",
    y     = NULL,
    color = "Year"
  )
#ggsave("suppfig2.png")

#write.csv(long_props, "long_props.csv")
#write.csv(summ_props, "summ_props.csv")

```

Supplemental Figure 3 - PCA of of standardized soil variables
```{r SF3}

#look at soil data as a whole 
#esquisser(data)

#for PCA will likely need to remove S site
#S is such a strong outlier for most micronutrients since was a historical mine site

soil_data <- data_filt %>% 
  filter(anon_id != "S")


#esquisser(soil_data)
#note, looking at this for both years together
#similar trends when looking at indiv years via esquisser


densityplot(soil_data$p_lb_a) 
soil_data$log_p <- log(soil_data$p_lb_a)
densityplot(soil_data$log_p) #better

densityplot(soil_data$mg_lb_a) 
soil_data$log_mg <- log(soil_data$mg_lb_a)
densityplot(soil_data$log_mg) #better

densityplot(soil_data$ca_lb_a) 
soil_data$log_ca <- log(soil_data$ca_lb_a)
densityplot(soil_data$log_ca) #better

densityplot(soil_data$k_lb_a) 
soil_data$log_k <- log(soil_data$k_lb_a)
densityplot(soil_data$log_k) #better

densityplot(soil_data$zn_ppm) 
soil_data$log_zn <- log(soil_data$zn_ppm)
densityplot(soil_data$log_zn) #better

densityplot(soil_data$fe_ppm) #already fine

densityplot(soil_data$mn_ppm) #already fine

densityplot(soil_data$b_ppm) 
soil_data$log_b <- log(soil_data$b_ppm)
densityplot(soil_data$log_b) #slightly better?

densityplot(soil_data$cu_ppm) 
soil_data$log_cu <- log(soil_data$cu_ppm)
densityplot(soil_data$log_cu) #better

densityplot(soil_data$so4_ppm) 
soil_data$log_so4 <- log(soil_data$so4_ppm)
densityplot(soil_data$log_so4) #better


densityplot(soil_data$cec_meq_100g) #fine

#will need to convert percent sand, silt, clay to composition object (since they are interdependent)

soil_texture <- soil_data[, c("sand", "silt", "clay")]
soil_texture_comp <- acomp(soil_texture)

#CLR transformation
soil_texture_clr <- clr(soil_texture_comp)

# Optional: check result
head(soil_texture_clr)

colnames(soil_texture_clr) <- c("sand_clr", "silt_clr", "clay_clr") #diff names so no merge issues

# 3. Combine with other variables for PCoA or modeling
soil_data_clr <- cbind(soil_data, as.data.frame(soil_texture_clr))

densityplot(soil_data_clr$sand_clr) #Fine
densityplot(soil_data_clr$silt_clr) #Fine
densityplot(soil_data_clr$clay_clr) #Fine



#only cols for PCA
soil_data_r <- soil_data_clr %>% 
  dplyr::select(stability,
         om_perc, 
         actinomycetes, am_fungi, #all PLFA data are very collinear, only putting these two in since they are of particular interest to the hypothesis
         ph, cec_meq_100g, #CEC may be redundant
         log_p, log_mg, log_ca, log_k,
         log_zn, log_b, log_cu, mn_ppm, fe_ppm,
         log_so4,
         sand_clr, silt_clr, clay_clr, 
         bulk_density, #Bulk Density was measured once in 2022, use same for 2023 PCA
         year
  ) %>% na.omit() #removes Q farm, since did not have some measurements taken only in 2023


# 3. Separate Predictors from Response
predictors22 <- soil_data_r %>%
  filter(year == "2022")

predictors22 <- predictors22 %>%
  dplyr::select( -year)

predictors23 <- soil_data_r %>%
  filter(year == "2023")

predictors23 <- predictors23 %>%
  dplyr::select( -year)

# A named vector where the 'name' is the original variable
# and the 'value' is the label you want to display
predictor_labels <- c(
  "stability" = "Stability",
  "cec_meq_100g"  = "CEC",
  "actinomycetes" = "Actino",
  "am_fungi"       = "AMF",
  "om_perc"        = "OM",
  "sand_clr" = "Sand",
  "silt_clr" = "Silt",
  "clay_clr" = "Clay",
  "log_p" = "P",
  "log_mg" = "Mg",
  "log_ca" = "Ca",
  "log_k" = "K",
  "log_zn" = "Zn",
  "log_b" = "B",
  "log_cu" = "Cu",
  "log_so4" = "SO4",
  "fe_ppm" = "Fe",
  "mn_ppm" = "Mn",
  "ph" = "ph",
  "bulk_density" = "Bulk Density"
)


#Run PCA with Centering & Scaling
# prcomp() automatically does mean-centering when center=TRUE,
# and divides by SD when scale.=TRUE


pca_result22 <- prcomp(predictors22, center = TRUE,  scale. = TRUE) #z score normalizes data
pca_result23 <- prcomp(predictors23, center = TRUE, scale. = TRUE)

#Extract PCA Scores (Samples) and Loadings (Variables)
pca_scores23 <- as.data.frame(pca_result23$x)       # PC coordinates of each sample
pca_scores23$stability <- predictors23$stability # Add soil stability for coloring

pca_scores22 <- as.data.frame(pca_result22$x)       # PC coordinates of each sample
pca_scores22$stability <- predictors22$stability # Add soil stability for coloring


pca_loadings22 <- as.data.frame(pca_result22$rotation)  # Loadings (arrows) for each variable
pca_loadings22$Variable <- rownames(pca_loadings22)
pca_loadings22$Label <- predictor_labels[pca_loadings22$Variable]

pca_loadings23 <- as.data.frame(pca_result23$rotation)  # Loadings (arrows) for each variable
pca_loadings23$Variable <- rownames(pca_loadings23)
pca_loadings23$Label <- predictor_labels[pca_loadings23$Variable]

#Compute Variance Explained by PC1 and PC2 (for plot labels)
var_explained22 <- pca_result22$sdev^2 / sum(pca_result22$sdev^2)
pc1_var22 <- round(var_explained22[1] * 100, 1)
pc2_var22 <- round(var_explained22[2] * 100, 1)


var_explained23 <- pca_result23$sdev^2 / sum(pca_result23$sdev^2)
pc1_var23 <- round(var_explained23[1] * 100, 1)
pc2_var23 <- round(var_explained23[2] * 100, 1)

#for only showing highest corr vectors
# Calculate correlation of each original variable with PC1 and PC2
# Get original variables (scaled and centered)
scaled_22 <- scale(predictors22)
scaled_23 <- scale(predictors23)

# Correlation of original variables with PCs (scores)
cor_PC1 <- abs(cor(scaled_22, pca_scores22$PC1))
cor_PC2 <- abs(cor(scaled_22, pca_scores22$PC2))

cor_PC1_23 <- abs(cor(scaled_23, pca_scores23$PC1))
cor_PC2_23 <- abs(cor(scaled_23, pca_scores23$PC2))
# Combine correlations and find the top ones

cor_matrix <- data.frame(Variable = rownames(cor_PC1), 
                         PC1 = cor_PC1[,1], 
                         PC2 = cor_PC2[,1])

cor_matrix23 <- data.frame(Variable = rownames(cor_PC1_23), 
                         PC1 = cor_PC1_23[,1], 
                         PC2 = cor_PC2_23[,1])

#cor not abs for discussion
cor_PC1_notabs <- cor(scaled_22, pca_scores22$PC1)
cor_PC2_notabs <- cor(scaled_22, pca_scores22$PC2)

cor_PC1_23_notabs <- cor(scaled_23, pca_scores23$PC1)
cor_PC2_23_notabs <- cor(scaled_23, pca_scores23$PC2)
# Combine correlations and find the top ones

cor_matrix_notabs <- data.frame(Variable = rownames(cor_PC1_notabs), 
                         PC1 = cor_PC1_notabs[,1], 
                         PC2 = cor_PC2_notabs[,1])

cor_matrix23_notabs <- data.frame(Variable = rownames(cor_PC1_23_notabs), 
                           PC1 = cor_PC1_23_notabs[,1], 
                           PC2 = cor_PC2_23_notabs[,1])

#wrote files

#write.csv(cor_matrix, "cor_matrix_22_pca.csv")
#write.csv(cor_matrix23, "cor_matrix_23_pca.csv")
#write.csv(cor_matrix_notabs, "cor_matrix_22_pca_notabs.csv")
#write.csv(cor_matrix23_notabs, "cor_matrix_23_pca_notabs.csv")

# Choose a threshold or top N variables
cor_matrix$MaxCorr <- pmax(cor_matrix$PC1, cor_matrix$PC2)
cor_matrix23$MaxCorr <- pmax(cor_matrix23$PC1, cor_matrix23$PC2)
# For example, keep top 16 variables by max correlation
top_vars <- head(cor_matrix[order(-cor_matrix$MaxCorr), ], 16)$Variable
top_vars23 <- head(cor_matrix23[order(-cor_matrix23$MaxCorr), ], 16)$Variable

# Filter pca_loadings22 to only keep top_vars
pca_loadings22_top <- subset(pca_loadings22, Variable %in% top_vars)
pca_loadings23_top <- subset(pca_loadings23, Variable %in% top_vars23)

#Plotting
#    - Points = samples, colored by 'stability'
#    - Arrows = loadings for each predictor variable
#    - geom_text_repel to label vectors


#2022

pca22 <- ggplot(pca_scores22, aes(x = PC1, y = PC2, color = stability)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_viridis()+
  #scale_color_gradient(low = "blue", high = "red") +
  geom_segment(
    data = pca_loadings22_top,
    aes(x = 0, y = 0, xend = PC1 * 8, yend = PC2 * 8),
    arrow = arrow(length = unit(0.3, "cm")), 
    color = "darkgrey",
    inherit.aes = FALSE
  ) +
  geom_text_repel(
    data = pca_loadings22_top,
    aes(x = PC1 * 8.4, y = PC2 * 8.4, label = Label),
    size = 3, fontface = "bold",
   # box.padding   = 0.6,  # how much space around the text box
    #point.padding = 0.8,  # how much space around the point
    inherit.aes = FALSE
  ) +
  theme_minimal(base_size = 12) +
  labs(
    title = "2022",
    x = paste0("PC1 (", pc1_var22, "%)"),
    y = paste0("PC2 (", pc2_var22, "%)"),
    color = "Stability"
  ) +
  theme(base.family = "Helvetica",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right",
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

pca22

#2023

pca23 <- ggplot(pca_scores23, aes(x = PC1, y = PC2, color = stability)) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_viridis()+
  #scale_color_gradient(low = "blue", high = "red") +
  geom_segment(
    data = pca_loadings23_top,
    aes(x = 0, y = 0, xend = PC1 * 8, yend = PC2 * 8),
    arrow = arrow(length = unit(0.3, "cm")), 
    color = "darkgrey",
    inherit.aes = FALSE
  ) +
  geom_text_repel(
    data = pca_loadings23_top,
    aes(x = PC1 * 8.2, y = PC2 * 8.2, label = Label),
    size = 3,
    fontface = "bold",
    nudge_x = 0, 
    nudge_y = 0,
    direction = "both",
    segment.color = NA,  # no line from label to point
    max.overlaps = Inf,
    box.padding = 0.15,   # smaller space around label box
    point.padding = 0.1,
    force = 2, 
    inherit.aes = FALSE
  ) +
  theme_minimal(base_size = 12) +
  labs(
    title = "2023",
    x = paste0("PC1 (", pc1_var23, "%)"),
    y = paste0("PC2 (", pc2_var23, "%)"),
    color = "Stability"
  ) +
  theme(base.family = "Helvetica",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right",
    plot.title = element_text(face = "bold", hjust = 0.5)
  )


pca_plot <- ggarrange(pca22,
          pca23,
          ncol = 2, nrow = 1,
          labels = c("A", "B"),
          common.legend = TRUE, legend = "right")
pca_plot



#ggsave("pca_soil_stability.png", pca_plot,  dpi = 600)
#ggsave("pca_soil_stability.eps", pca_plot,dpi = 600)
#ggsave("pca_soil_stability.tiff", pca_plot,  dpi = 600)
#ggsave("pca_soil_stability.svg", pca_plot, dpi = 600)

#save PCA scores
#write.csv(pca_scores22, "pca_scores22.csv")
#write.csv(pca_scores23, "pca_scores23.csv")
#write reduced data
#write.csv(soil_data_r, "soil_data_red_PCA.csv")

```

Figure 1 - Boxplots and scatterplot, some statistics
```{r F1}

boxplot.data <- data_filt %>%
  pivot_longer(cols = c("actinomycetes", "am_fungi", "gram_negative", "gram_positive", "eukaryote", "fungi"), 
               names_to = "PLFA_Group", values_to = "nmol.p.g.soil")

# Format factors 
boxplot.data$PLFA_Group <- as.factor(boxplot.data$PLFA_Group)
boxplot.data$year <- factor(boxplot.data$year)

# Remove rows with NA in key columns
boxplot.cleaned <- boxplot.data %>%
  filter(!is.na(nmol.p.g.soil),
         !is.na(PLFA_Group),
         !is.na(year),
         !is.na(anon_id))

# Rename and reorder PLFA groups
boxplot.cleaned <- boxplot.cleaned %>%
  mutate(PLFA_Group = fct_recode(PLFA_Group,
                                 "Actinobacteria" = "actinomycetes",
                                 "Gram-Negative\n Bacteria" = "gram_negative",
                                 "AM Fungi" = "am_fungi",
                                 "Gram-Positive\n Bacteria" = "gram_positive",
                                 "Eukaryote" = "eukaryote",
                                 "Fungi" = "fungi"
  )) %>%
  mutate(PLFA_Group = fct_reorder(
    PLFA_Group,
    nmol.p.g.soil,
    .fun = median,
    .desc = TRUE
  ))

# Test year effect within each PLFA group
get_stars <- function(p) {
  ifelse(p < 0.001, "***",
         ifelse(p < 0.01, "**",
                ifelse(p < 0.05, "*", "")))
}

# Perform Wilcoxon test for each group
year_effect_stats <- boxplot.cleaned %>%
  group_by(PLFA_Group) %>%
  summarise(
    p_value = wilcox.test(nmol.p.g.soil ~ year)$p.value,
    .groups = "drop"
  ) %>%
  mutate(stars = sapply(p_value, get_stars))

# Calculate y positions for significance labels
y_positions <- boxplot.cleaned %>%
  group_by(PLFA_Group) %>%
  summarise(y_max = max(nmol.p.g.soil, na.rm = TRUE), .groups = "drop") %>%
  mutate(y_label = y_max + 0.08 * max(y_max))

# Merge stars with positions
sig_labels <- year_effect_stats %>%
  left_join(y_positions, by = "PLFA_Group") %>%
  filter(stars != "")

# Create the plot
plfa_plot <- ggplot(boxplot.cleaned, aes(x = PLFA_Group, 
                                         y = nmol.p.g.soil, fill = year)) +
  geom_boxplot(alpha = 0.8) +
  scale_fill_manual(values = c("2022"="#1f77b4","2023"="darkorange")) +
  labs(y = "nmol PLFA per g soil", x = "PLFA Group", fill = "Year") +
  theme_classic(base_family = "Helvetica") +
  theme(
    legend.position = "top",
    text = element_text(size = 14)
  )

# Add significance stars if any groups show year effects
if(nrow(sig_labels) > 0) {
  plfa_plot <- plfa_plot +
    geom_text(data = sig_labels,
              aes(x = PLFA_Group, y = y_label, label = stars),
              inherit.aes = FALSE,
              size = 5, vjust = 0, fontface = "bold", color = "black")
}

plfa_plot

#ggsave("boxplot.png")
#ggsave("boxplot.tiff")
#ggsave("boxplot.eps")
#ggsave("boxplot.svg")

#write.csv(boxplot.cleaned, "boxplot.cleaned.csv")

# Function to convert p-values to cap at ***
format_p_signif <- function(p) {
  if (p < 0.001) return("***")
  else if (p < 0.01) return("**")
  else if (p < 0.05) return("*")
  else return("ns")
}

# OM plot; Figure 1B
om_result <- wilcox.test(om_perc ~ year, data = data_filt)
om_p_label <- format_p_signif(om_result$p.value)

om_plot <- ggplot(data_filt, aes(x = year, y = om_perc, fill = year)) +
  geom_boxplot(width = 0.5, alpha = 0.8) +
  scale_fill_manual(values = c("2022"="#1f77b4","2023"="darkorange")) +
  annotate("text", x = 1.5, y = max(data_filt$om_perc, na.rm = TRUE) * 1.001,
           label = om_p_label, size = 5, fontface = "bold") +
  labs(y = "Organic Matter (%)", x = "") +
  theme_classic(base_family = "Helvetica") +
  theme(
    legend.position = "none",
    text = element_text(size = 10),
    plot.title = element_text(margin = margin(b = 5))
  )
om_plot

# Stability Plot Fig 1 A
stab_result <- wilcox.test(stability ~ year, data = data_filt)
stab_p_label <- format_p_signif(stab_result$p.value)

stability_plot <- ggplot(data_filt, aes(x = year, y = stability, fill = year)) +
  geom_boxplot(width = 0.5, alpha = 0.8) +
  scale_fill_manual(values = c("2022"="#1f77b4","2023"="darkorange")) +
  annotate("text", x = 1.5, y = max(data_filt$stability, na.rm = TRUE) * 1.001,
           label = stab_p_label, size = 5, fontface = "bold") +
  labs(y = "Macroaggregate Stability Index", x = "") +
  theme_classic(base_family = "Helvetica") +
  theme(
    legend.position = "none",
    text = element_text(size = 10),
    plot.title = element_text(margin = margin(b = 5))
  )
stability_plot

# Stability by OM plot; Figure 1C

data.filt22 <- data_filt %>%
  filter(year == 2022)

data.filt23 <- data_filt %>%
  filter(year == 2023)

stability22 <- lm(stability ~ om_perc, data = data.filt22)
stab_r2_22 <- summary(stability22)$adj.r.squared
stab_r2_22 #0.4416

stability23 <- lm(stability ~ om_perc, data = data.filt23)
stab_r2_23 <- summary(stability23)$adj.r.squared
stab_r2_23 #0.3199

# Create labels for each R²
st_label_spring <- paste0("2022 Adj.R² = ", round(stab_r2_22, 2))
st_label_fall   <- paste0("2023 Adj.R² = ", round(stab_r2_23, 2))

stbXom.plot <- ggplot(data = data_filt, aes(x = om_perc, y = stability, color = as.factor(year))) +
  geom_point(alpha = 0.8) +
  geom_smooth(data = data_filt,
              aes(x = om_perc, y = stability, color = as.factor(year)),
              method = "lm",
              se = FALSE,
              inherit.aes = FALSE) +
  scale_size_continuous(range = c(1, 5.5)) +
  scale_color_manual(values = c("2022"="#1f77b4","2023"="darkorange")) +
  labs(x = "Organic Matter (%)", y = "Macroaggregate Stability Index", color = element_blank()) +
  annotate("text", x = 3.5, y = 2.3, label = st_label_spring, color = "#1f77b4", size = 3.5) +
  annotate("text", x = 3.5, y = 2.1, label = st_label_fall, color = "darkorange", size = 3.5) +
  theme_classic(base_size = 9, base_family = "Helvetica") +
  theme(
    legend.position = "none",
    text = element_text(size = 10),
    plot.title = element_text(margin = margin(b = 5))
  )

stbXom.plot

# Combine using patchwork
# Figure 1
Figure1_All <- ((stability_plot | om_plot | stbXom.plot) / plfa_plot) +
  plot_layout(heights = c(1.3, 2)) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 14, face = "bold"))

((stability_plot | om_plot | stbXom.plot) / plfa_plot) +
  plot_layout(heights = c(1.3, 2)) +
  plot_annotation(tag_levels = 'A')&
  theme(plot.tag = element_text(size = 14, face = "bold"))

ggsave(filename = "fig1.png", plot = Figure1_All)
ggsave(filename = "fig1.tiff", plot = Figure1_All)
ggsave(filename = "fig1.pdf", plot = Figure1_All)

# % increases for text

# 1. Overall % change between year-means
year_means <- boxplot.data %>%
  filter(!is.na(stability), !is.na(om_perc), !is.na(total_plfa_nmol_p_g_soil)) %>%
  group_by(year) %>%
  summarise(across(
    c(stability, om_perc, total_plfa_nmol_p_g_soil),
    ~ mean(.x, na.rm = TRUE)
  ))

# Compute overall pct change (2023 vs 2022)
overall_pct <- year_means %>%
  pivot_wider(names_from = year, values_from = c(stability, om_perc, total_plfa_nmol_p_g_soil)) %>%
  transmute(
    stability_pct   = (stability_2023 - stability_2022) / abs(stability_2022) * 100,
    om_perc_pct      = (om_perc_2023   - om_perc_2022)   / om_perc_2022   * 100,
    TotalPLFA_pct   = (total_plfa_nmol_p_g_soil_2023 - total_plfa_nmol_p_g_soil_2022) /
      total_plfa_nmol_p_g_soil_2022 * 100
  )
overall_pct

# For PLFA groups
# 1. Grand-mean % change per PLFA group
group_means <- boxplot.data %>%
  filter(!is.na(nmol.p.g.soil)) %>%
  group_by(PLFA_Group, year) %>%
  summarise(mean_abund = mean(nmol.p.g.soil, na.rm = TRUE), .groups="drop")

group_pct_change <- group_means %>%
  pivot_wider(names_from = year, values_from = mean_abund, 
              names_prefix = "mean_") %>%
  mutate(pct_change = (mean_2023 - mean_2022) / mean_2022 * 100) %>%
  arrange(desc(pct_change))

# 2. Average of per-plot % change per PLFA group
per_plot_group <- boxplot.data %>%
  filter(!is.na(nmol.p.g.soil)) %>%
  group_by(anon_id, PLFA_Group, year) %>%
  summarise(plot_mean = mean(nmol.p.g.soil, na.rm = TRUE), .groups="drop") %>%
  pivot_wider(names_from = year, values_from = plot_mean, 
              names_prefix = "plot_") %>%
  mutate(pct_change = (plot_2023 - plot_2022) / plot_2022 * 100) 

group_plot_pct <- per_plot_group %>%
  group_by(PLFA_Group) %>%
  summarise(mean_pct_change = mean(pct_change, na.rm = TRUE),
            sd_pct_change   = sd(pct_change,   na.rm = TRUE),
            n_plots         = sum(!is.na(pct_change)), .groups="drop") %>%
  arrange(desc(mean_pct_change))

# View results
group_pct_change
group_plot_pct

# PLFA group relative abundances for text
# Using total assigned PLFAs (not including 25% ish of unassigned PLFAs)
rel_summary <- boxplot.data %>%
  # 1. Calculate each group's % of its plot's total PLFA
  group_by(anon_id, year) %>%
  mutate(
    total_PLFA = sum(nmol.p.g.soil, na.rm = TRUE),
    rel_abund  = nmol.p.g.soil / total_PLFA * 100
  ) %>%
  ungroup() %>%
  
  # 2. Mean relative abundance by group & year
  group_by(PLFA_Group, year) %>%
  summarise(
    mean_rel = mean(rel_abund, na.rm = TRUE),
    sd_rel   = sd(rel_abund,   na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  
  # 3. Pivot to wide with clear names
  pivot_wider(
    names_from  = year,
    values_from = c(mean_rel, sd_rel),
    names_sep   = "_"
  ) %>%
  
  # 4. Compute the average of the two year-means
  mutate(
    mean_rel_both = (mean_rel_2022 + mean_rel_2023) / 2
  ) %>%
  arrange(desc(mean_rel_both))

rel_summary

# Combined AMR, euk, fungi
# 5.47 + 2.83 + 2.37 = 10.67


```

Figure 3 - Ridge regression and coefficient ridgeplots
```{r F3}
set.seed(2026)

# Function for bootstrap ridge regression
bootstrap_ridge <- function(X, y, n_boot = 1000, alpha_val = 0) {
  
  # First, fit one model to get correct variable names
  fit0 <- cv.glmnet(X, y, alpha = alpha_val)
  best_lambda0 <- fit0$lambda.min
  
  # Safe way to extract coefficients excluding intercept
  coef_vector0 <- coef(fit0, s = best_lambda0)
  coef_vector0 <- coef_vector0[setdiff(rownames(coef_vector0), "(Intercept)"), , drop = FALSE]
  
  coef_names <- rownames(coef_vector0)
  
  # Initialize coefficient matrix with correct names
  coefs <- matrix(NA, nrow = n_boot, ncol = length(coef_names))
  colnames(coefs) <- coef_names
  
  # Run bootstrap loop
  for (i in 1:n_boot) {
    idx <- sample(1:nrow(X), replace = TRUE)
    X_resample <- X[idx, ]
    y_resample <- y[idx]
    
    fit <- cv.glmnet(X_resample, y_resample, alpha = alpha_val)
    best_lambda <- fit$lambda.min
    
    coef_vector <- coef(fit, s = best_lambda)
    coef_vector <- coef_vector[setdiff(rownames(coef_vector), "(Intercept)"), , drop = FALSE]
    
    # Defensive check to ensure dimensions match
    coef_values <- as.numeric(coef_vector)
    names(coef_values) <- rownames(coef_vector)
    
    # Match current coefficients to full list — fill missing with 0 if needed
    coef_full <- rep(0, length(coef_names))
    names(coef_full) <- coef_names
    coef_full[names(coef_values)] <- coef_values
    
    coefs[i, ] <- coef_full
  }
  
  # Create dataframe
  coef_df <- as.data.frame(coefs)
  
  # Summary table
  coef_summary <- coef_df %>%
    pivot_longer(cols = everything(), names_to = "PLFA_Group", values_to = "Coefficient") %>%
    group_by(PLFA_Group) %>%
    summarise(Mean = mean(Coefficient),
              Median = median(Coefficient),
              SD = sd(Coefficient),
              .groups = "drop")
  
  # Optional sanity check
  cat("\nBootstrap ridge sanity check — variable names:\n")
  print(colnames(coef_df))
  
  cat("\nExample of first row of coefficients:\n")
  print(coef_df[1, ])
  
  return(list(summary = coef_summary, raw = coef_df))
}




# Function to calculate performance over repeated train/test splits
ridge_performance <- function(X, y, n_splits = 50, seed = 2025) { #should seed match seed above?)
  set.seed(seed)
  test_r2 <- numeric(n_splits)
  test_mse <- numeric(n_splits)
  cv_mse <- numeric(n_splits)

  for (i in 1:n_splits) {
    idx <- sample(seq_len(nrow(X)), size = 0.7 * nrow(X))
    X_train <- X[idx, ]
    y_train <- y[idx]
    X_test  <- X[-idx, ]
    y_test  <- y[-idx]

    cv_fit <- cv.glmnet(X_train, y_train, alpha = 0)
    lambda_min <- cv_fit$lambda.min

    preds <- predict(cv_fit, newx = X_test, s = lambda_min)
    resid <- y_test - preds

    test_r2[i]  <- 1 - sum(resid^2) / sum((y_test - mean(y_test))^2)
    test_mse[i] <- mean(resid^2)
    cv_mse[i]   <- min(cv_fit$cvm)
  }

  return(list(
    median_r2 = median(test_r2),
    mean_mse = mean(test_mse),
    mean_cv_mse = mean(cv_mse),
    all_r2 = test_r2,
    all_mse = test_mse
  ))
}


#nature esque plotting function
theme_nature <- function(base_size = 12, base_family = "Helvetica") {
  theme_minimal(base_size = base_size, base_family = base_family) +
    theme(
      axis.text = element_text(color = "black"),
      axis.title.x = element_text( size = base_size), #hjust = 0.38,vjust = -1
      axis.title.y = element_text(size = base_size),
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_line(color = "grey90", linewidth = 0.3),
      plot.margin = margin(5, 15, 10, 5),
      legend.position = "top",
      legend.title = element_blank(),
      legend.key.height = unit(0.5, "lines"),
      legend.text = element_text(size = base_size)
    )
}



#Remove any rows with missing values in the relevant columns
#     This ensures the model has complete cases.
#using no outlier data, and log trans Fungi. Euk who tf knows

data_lm22 <- na.omit(data_2022[, c("stability", 
                      "actinomycetes", "am_fungi", 
                      "gram_negative", "gram_positive", 
                      "log_fungi", "log_eukaryote", 
                      "om_perc")])

data_lm23 <- na.omit(data_2023[, c("stability", 
                                   "actinomycetes", "am_fungi", 
                      "gram_negative", "gram_positive", 
                      "log_fungi", "log_eukaryote", 
                      "om_perc")])



data_lm22_scaled <- data_lm22 %>%
  mutate(across(c(stability, actinomycetes, am_fungi, gram_positive, gram_negative, 
                  log_fungi, log_eukaryote, om_perc), ~ as.numeric(scale(.))))  # z-score normalize OMperc and stability, didn't have stability scaled before.
data_lm23_scaled <- data_lm23 %>%
  mutate(across(c(stability, actinomycetes, am_fungi, gram_positive, gram_negative, 
                  log_fungi, log_eukaryote, om_perc), ~ as.numeric(scale(.)))) 

summary(data_lm22_scaled) #mean = 0
sd(data_lm22_scaled$am_fungi) #sd = 1
summary(data_lm23_scaled) #mean = 0
sd(data_lm23_scaled$am_fungi) #sd = 1


#      A rule of thumb is VIF > 5 (some say >10) suggests problematic multicollinearity.
#      High VIF means that predictor is highly correlated with other predictors.
#So Actino, AMF, GN, GP, are all HIGHLY collin. for both years

#esquisser(data_lm23)

full_model22 <- lm(stability ~ am_fungi + gram_negative + 
                   gram_positive + actinomycetes + log_eukaryote + log_fungi +
                   om_perc, data = data_lm22_scaled) 

summary(full_model22) # only OMperc sig
par(mfrow = c(2, 2))
plot(full_model22)

vif_values22 <- vif(full_model22)
vif_values22 

#AMFungi  GramNegative  GramPositive Actinomycetes log_Eukaryote     log_Fungi        OMPerc 
#15.201221     61.098961     46.278824     25.779458      2.773259      1.899392      3.360675 

full_model23 <- lm(stability ~ am_fungi + gram_negative + 
                   gram_positive + actinomycetes + log_eukaryote + log_fungi +
                   om_perc, data = data_lm23_scaled) 

summary(full_model23) # only OMperc sig
par(mfrow = c(2, 2))
plot(full_model23) 

vif_values23 <- vif(full_model23)
vif_values23 
#AMFungi  GramNegative  GramPositive Actinomycetes log_Eukaryote     log_Fungi        OMPerc 
#12.919141     35.468566     37.882851     26.561784      2.057834      1.862385      3.820509 


#write.csv(data_lm22, "data_matrix_2022.csv")
#write.csv(data_lm23, "data_matrix_2023.csv")
#write.csv(data_lm22_scaled, "norm_data_matrix_2022.csv")
#write.csv(data_lm23_scaled, "norm_data_matrix_2023.csv")




# 2022 data
X_22 <- model.matrix(stability ~ actinomycetes + am_fungi + gram_positive + 
                       gram_negative + log_eukaryote + om_perc + log_fungi,  
                     data = data_lm22_scaled)[, -1]
y_22 <- data_lm22_scaled$stability

# 2023 data
X_23 <- model.matrix(stability ~ actinomycetes + am_fungi + gram_positive + 
                       gram_negative + log_eukaryote + om_perc + log_fungi,  
                     data = data_lm23_scaled)[, -1]
y_23 <- data_lm23_scaled$stability


ridge_out22 <- bootstrap_ridge(X_22, y_22)
ridge_out23 <- bootstrap_ridge(X_23, y_23)

ridge_boot22 <- ridge_out22$summary
ridge_boot23 <- ridge_out23$summary

raw_22 <- ridge_out22$raw %>% mutate(year = "2022")
raw_23 <- ridge_out23$raw %>% mutate(year = "2023")


raw_all <- bind_rows(raw_22, raw_23) %>%
  pivot_longer(cols = -year, names_to = "PLFA_Group", values_to = "Coefficient")


#model performance
perf_2022 <- ridge_performance(X_22, y_22)
perf_2023 <- ridge_performance(X_23, y_23)

perf_2022$median_r2  #0.43
perf_2022$mean_cv_mse  #0.59

perf_2023$median_r2  #0.28
perf_2023$mean_cv_mse  #0.66


#visualize distributions

ggplot(raw_all, aes(x = Coefficient, color = year)) +
  geom_density(alpha = 0.4) +
  facet_wrap(~ PLFA_Group, scales = "free") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal(base_size = 12) +
  labs(title = "Bootstrap Coefficient Distributions",
       x = "Bootstrapped Coefficient Value", y = "Density")


#combine and order

ridge_ci <- raw_all %>%
  group_by(PLFA_Group, year) %>%
  summarise(Median = median(Coefficient),
            Lower = quantile(Coefficient, 0.025), #90% is 0.05 and 0.95 CI
            Upper = quantile(Coefficient, 0.975),
            .groups = "drop")



ridge_ci <- ridge_ci %>%
  mutate(PLFA_Group = fct_recode(PLFA_Group,
                                 "Organic Matter" = "om_perc",
                                 "Actinobacteria" = "actinomycetes",
                                 "Gram-Negative\nBacteria" = "gram_negative",
                                 "AM Fungi" = "am_fungi",
                                 "Eukaryote" = "log_eukaryote",
                                 "Gram-Positive\nBacteria" = "gram_positive",
                                 "Fungi" = "log_fungi"))

raw_all <- raw_all %>%
  mutate(PLFA_Group = fct_recode(PLFA_Group,
                                 "Organic Matter" = "om_perc",
                                 "Actinobacteria" = "actinomycetes",
                                 "Gram-Negative\nBacteria" = "gram_negative",
                                 "AM Fungi" = "am_fungi",
                                 "Eukaryote" = "log_eukaryote",
                                 "Gram-Positive\nBacteria" = "gram_positive",
                                 "Fungi" = "log_fungi"))

#Define ordered_groups AFTER recoding
ordered_groups <- ridge_ci %>%
  filter(year == "2022") %>%
  arrange(Median) %>%
  pull(PLFA_Group) %>% as.character()

#Apply ordered factor levels to both ridge_ci and raw_all
ridge_ci$PLFA_Group <- factor(ridge_ci$PLFA_Group, levels = ordered_groups)
raw_all$PLFA_Group  <- factor(raw_all$PLFA_Group,  levels = ordered_groups)


# Define this y position once, so both use it
ridge_ci <- ridge_ci %>%
  mutate(y_pos = as.numeric(fct_reorder(PLFA_Group, Median)) +
           ifelse(year == "2022", -0.1, 0.1))

#make 2022 first in legend
raw_all$year <- factor(raw_all$year, levels = c("2022", "2023"))

#plot
ggplot(raw_all, aes(x = Coefficient, y = PLFA_Group, fill = year)) +
  geom_density_ridges(alpha = 0.6, scale = 0.8, rel_min_height = 0.01, color = "grey20") +
  # Add CI as segment at y position — approximate y position by year dodge
  geom_segment(data = ridge_ci,
               aes(x = Lower, xend = Upper,
                   y = y_pos, yend = y_pos,
                   color = "grey"),
               linewidth = 0.5, show.legend = FALSE) +
  # Median point
  geom_point(data = ridge_ci,
             aes(x = Median, y = y_pos, fill = year),
             shape = 21, color = "black", size = 3, stroke = 0.75, show.legend = FALSE) +
  #0 line
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  labs(
       x = "Standardized Ridge Coefficient\n(Median ± 95% CI)",
       y = "PLFA Group") +
  scale_fill_manual(values = c("2022" = "#1f77b4", "2023" = "darkorange")) +
  scale_color_manual(values = c("2022" = "#1f77b4", "2023" = "darkorange")) +
  scale_x_continuous(limits = c(-0.8, 0.8)) +
  theme_nature()+
  theme(axis.title.y = element_text(margin = margin(r = 3)),  # reduce right margin
axis.text.y  = element_text(margin = margin(r = -50)) # tighter label-to-axis spacing
)

coeff_ridge_avg <- raw_all %>%
  group_by(PLFA_Group, year) %>%
  summarise(mean = mean(Coefficient),
            median = median(Coefficient))

write.csv(coeff_ridge_avg, "coeff_ridge_avg.csv")
# Use ggsave to export
#ggsave("Figure3_Ridge.pdf", width = 85, height = 120, units = "mm", dpi = 300, device = cairo_pdf)

#ggsave("Figure3_Ridge.tiff", width = 85, height = 120, units = "mm", dpi = 300, compression = "lzw")

#ggsave("Figure3_Ridge.png")


#write.csv(raw_all, "raw_all.csv")


```

Supp Figure 4 - Proportional PLFA Boxplot and ART-ANOVA

```{r SFig4}
#Proportional Abundance Boxplots and Test

# Calculate total PLFA per sample for proportional abundance
# First, get the total for each sample (sum across all PLFA groups)
total_plfa_per_sample <- boxplot.data %>%
  group_by(index, anon_id, year) %>%
  summarise(total_plfa = sum(nmol.p.g.soil, na.rm = TRUE), .groups = 'drop')

# Add total PLFA to the dataset and calculate proportional abundance
boxplot.data <- boxplot.data %>%
  left_join(total_plfa_per_sample, by = c("index", "anon_id","year")) %>%
  mutate(prop_abundance = nmol.p.g.soil / total_plfa)

# Remove rows with NA in key columns
boxplot.cleaned <- boxplot.data %>%
  filter(!is.na(prop_abundance),
         !is.na(PLFA_Group),
         !is.na(year),
         !is.na(anon_id))

# Reorder based on descending value (using proportional abundance)
boxplot.cleaned <- boxplot.cleaned %>%
  mutate(PLFA_Group = fct_reorder(
    PLFA_Group,
    prop_abundance,
    .fun    = median,
    .desc   = TRUE
  ))

# Rename PLFA groups
boxplot.cleaned <- boxplot.cleaned %>%
  mutate(PLFA_Group = fct_recode(PLFA_Group,
                                 "Actinobacteria" = "actinomycetes",
                                 "Gram-Negative\nBacteria" = "gram_negative",
                                 "AM Fungi" = "am_fungi",
                                 "Eukaryote" = "eukaryote",
                                 "Gram-Positive\nBacteria" = "gram_positive",
                                 "Fungi" = "fungi"))

# Test year effect within each PLFA group
get_stars <- function(p) {
  ifelse(p < 0.001, "***",
         ifelse(p < 0.01, "**",
                ifelse(p < 0.05, "*", "")))
}

# Perform Wilcoxon test for each group
year_effect_stats_p <- boxplot.cleaned %>%
  group_by(PLFA_Group) %>%
  summarise(
    p_value = wilcox.test(prop_abundance ~ year)$p.value,
    .groups = "drop"
  ) %>%
  mutate(stars = sapply(p_value, get_stars))

# Calculate y positions for significance labels
y_positions_p <- boxplot.cleaned %>%
  group_by(PLFA_Group) %>%
  summarise(y_max = max(prop_abundance, na.rm = TRUE), .groups = "drop") %>%
  mutate(y_label = y_max + 0.08 * max(y_max))

# Merge stars with positions
sig_labels_p <- year_effect_stats_p %>%
  left_join(y_positions_p, by = "PLFA_Group") %>%
  filter(stars != "")

# Create the plot
plfa_plot_p <- ggplot(boxplot.cleaned, aes(x = PLFA_Group, 
                                           y = prop_abundance, fill = year)) +
  geom_boxplot(alpha = 0.8) +
  scale_fill_manual(values = c("2022"="#1f77b4","2023"="darkorange")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(y = "Proportional PLFA Abundance (%)", x = "PLFA Group", fill = "Year") +
  theme_classic(base_family = "Helvetica") +
  theme(
    legend.position = "top",
    text = element_text(size = 14)
  )

# Add significance stars if any groups show year effects
if(nrow(sig_labels_p) > 0) {
  plfa_plot_p <- plfa_plot_p +
    geom_text(data = sig_labels_p,
              aes(x = PLFA_Group, y = y_label, label = stars),
              inherit.aes = FALSE,
              size = 5, vjust = 0, fontface = "bold", color = "black")
}

plfa_plot_p

# Calculate median percent change for groups with significant year effects
median_by_year_group <- boxplot.cleaned %>%
  group_by(PLFA_Group, year) %>%
  summarise(median_prop = median(prop_abundance, na.rm = TRUE), .groups = "drop")

# Pivot to compare years
median_change <- median_by_year_group %>%
  pivot_wider(names_from = year, values_from = median_prop, names_prefix = "median_") %>%
  mutate(
    pct_change = (median_2023 - median_2022) * 100,
    direction = ifelse(pct_change > 0, "increase", "decrease")
  ) %>%
  # Join with significance results
  left_join(year_effect_stats_p, by = "PLFA_Group") %>%
  # Only show groups with significant year effects
  filter(stars != "") %>%
  arrange(desc(abs(pct_change)))

cat("\n=== Median Proportional Abundance Change (2022 to 2023) ===\n")
print(median_change[, c("PLFA_Group", "median_2022", "median_2023", "pct_change", "direction", "stars")])

# Write to CSV if desired
# write.csv(median_change, "prop_abundance_median_change.csv", row.names = FALSE)

ggsave("prop_plfa_boxplot.jpeg")
ggsave("prop_plfa_boxplot.tiff")
ggsave("prop_plfa_boxplot.png")
ggsave("prop_plfa_boxplot.pdf")

```

